Distribution of portfolio credit risk and capital savings

### 2019-08-14 Analysis/Draft/Draft.Rmd
Draft version with first try of IV and dif-in-dif analysis, plus decomposition plots

library(tidyverse)           ## Data manipulation, pipe operator
library(splitstackshape)     ## Command to expand data
library(data.table)          ## Command to bind lists
library(ggridges)            ## Density ridges plot
library(stargazer)           ## Latex tables
library(lfe)                 ## High-dimensional fixed effects
library(gridExtra)           ## Plot side-by-side
library(DescTools)           ## Command to winsorize variables

p_w <- c(0.0, 1)
```

```{r load data, include=FALSE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

@@

```
# Introduction

Many studies have questioned the reliability of banks' calculations of risk-weighted assets (RWA) for prudential purpose. The significant divergences found across banks are taken as indicating excessive subjectivity in the current rules governning banks' risk measurement and capital requirement calculations. Importantly, this subjectivity is unintended since differences in RWA driven by it do not reflect differences in risk. This paper shows that the dispersion _within_ banks' portfolio credit risk, more specifically the probability of default (PD) parameter, explains substantially the variation in RWA _across_ banks.

paragraph connecting to previous literature
Basel II introduced a concave mapping from probabilities of default to capital requiremets. As a consequence, banks can reduce requirements or increase risk-taking by widening their portfolio credit risk distribution.

To calculate equation X, or any other measure of capital savings arising from the concavity of the regulatory formula,  one needs data on individual, or group of assets' regulatory risk parameters. However, such granularity is not available in the common commercial balance sheet datasets, and, although very detailed, datasets from credit registers are restricted to one country and are not harmonized. Therefore, to analyze the drivers and implications of PD dispersion, I collected information on portfolio distribution from banks' individual Pillar-3 reports.\footnote{More specifically, I collected the information contained in tables that follow the CR5, CR6, CR10, CCR3, and CCR4 templates from the BCBS revised version of the Pillar 3 framework, or equivalent.} The sample consists of 50 large banking groups, from 10 countries, that have adopted the IRB approach. For each bank, I collected group consolidated information from the year the IRB approach was approved, which is early as 2007, until 2018. The dataset distinguishes between wholesale, retail and equity portfolios and, in most cases, between their sub-categories.\footnote{The sub-categories for wholesale are, corporate, banks, sovereign, and specialized lending, for retail, real estate, qualifying revolving credit, and other retail, and for equity, portfolios under the PD/LGD approach, and under the simple approach.} This level of portfolio breakdown is important because the regulatory formula varies across categories. It also allows me to control for variation in capital requirements due to business mix (sovereign portfolios require less capital than corporate) that could  be otherwise captured by my capital savings measure. Information on the share of portfolio under the standardized approach is also gathered, which allow me to control for the roll-out effect, i.e., strategical reduction in capital requirements by moving specific portfolios from the standardized to the IRB approach. Finally

paragraph on results

# Hypotheses and Data
In this section I describe the hypotheses of the paper by explaining the Basel II regulatory formulae under the IRB approach to calculate RWA for credit risk.

## Capital savings measure

@@

\begin{equation}
\text{Capital savings}_{i}= f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i),
\end{equation}

@@

## Data

Figure 1 shows that the log of _Capital savings_ varies considerably across banks. While most canadian banks seem to have not explore this possilibity to save capital, the european banks in our sample appear to have more than halved their capital requirement due to the distribution mechanism. Importantly, this comparison does not means that european banks are riskier than their canadian counterparties but only that their credit risk distribution is more disperse. The figure also shows that there capital savings have changed in time. However, the figure is not appropriated to see how this variable evolved in time.

```{r}
fig1_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
```

```{r 1, fig.height = 4, fig.width = 5,results="hide",fig.cap= fig1_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(name)) %>%
 ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
                          y = RW.s,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital savings") +

@@

Figure 2 shows that both _Capital savings_ and total capital ratio have increased slighly on average in time. After a spike in 2010, the average return on equity appears to have remained constant in time.

```{r}
fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE. The solid line is the average across banks for each year. The first and third quartiles are shown in the gray area.  "
```

```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth"}
mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
  dat <- data.frame(y = mean(x, na.rm = TRUE),
                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
  return(dat)
}

plot1 <- bank.data %>% 
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
 geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Capital savings') +
  theme_bw()

plot2 <- bank.data %>% mutate(tier1ratio= tier1ratio/100) %>% filter(year>2001)%>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Total capital ratio') +
  theme_bw()

plot3 <- bank.data %>% mutate(ROE = ROE/100) %>% filter(year>2001)%>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'ROE') +
  theme_bw()

@@

Figure 3 shows the linear relationship between our variable of interest, _Capital savings_ and three potential outcome variables: core tier 1 capital ratio, return on equity and, a measure of riskness, Z-score.

```{r, fig.height = 10, fig.width = 7, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
pillar3.expanded %>% 
  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
                       scale = 2 , from = 0, to = 0.05) +
  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    x = 'Probability of default') +
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())
```

Finally, we supplement the pillar-3 report data with annual information on banks' balance sheets obtained from the Bureau van Dijk's BankFocus database. Table 1 show summary statistics of the final dataset. The sample includes 25 banks. We restrict the analysis to banks for which we are able to collect information from Pillar-III reports. 
```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
time.decomposition %>% 
  select(time,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = time,
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  labs(x=NULL)

banks. We restrict the analysis to banks for which we are able to collect information from Pillar-III reports. 
```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
year.decomposition %>% 
  select(year,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = year,
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  labs(x=NULL)

The current geographical composition of the sample is: 1 norwegian, 1 italian, 1 spanish, 2 british, 4 australian, 6 canadian, and 10 american banks. The target geographical composition is: 1 norwegian, 1 austrian, 2 finish, 3 belgium, 3 danish, 4 italian, 4 spanish banks, 4 dutch, 4 australian, 4 swedish, 5 french, 5 british, 6 canadian, 7 german, and 10 american banks. This sums up to 63 IRB banks. The selection criteria is asset size.
```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
cross.section.decomposition %>% 
  mutate(order.s = abs(dispersion.gain))%>%
  select(name, Country, RW,order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = reorder(name, order.s),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  labs(x=NULL) +
    coord_flip()

```{r, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
cross.section.decomposition %>% 
  mutate(order.s = abs(dispersion.gain))%>%
  select(name, Country, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(value = abs(value),
         condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = reorder(name, order.s),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity',position = "fill")+
  labs(x=NULL) +
    coord_flip()

```{r,  fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
fig1_cap = "The figure plots the linear relationship between the capital savings measure and,in descending order, the tier 1 capital ratio, the return on equity using PL before taxes, and the Z-score. In each graph, a point is the average across time for each bank."
```

```{r relation, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
 bank.data %>% 
  select(name,RW.s, tier1ratio, tier.hat) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)%>%
  filter(RW.s<0.4)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="CT1 capital ratio") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()
```

```{r relation, fig.height = 4, fig.width = 5, results="hide",fig.cap= fig1_cap,out.width = "1\\textwidth"}
bank.data %>% 
  select(name,RW.s, totalcapitalratio, CAR.hat) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
  filter(RW.s<0.4)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="CT1 capital ratio") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()

```
Banks that only implement the standardized approach (SA) for credit risk most of the sample period will also be included in the dataset. The current geographical composition of SA banks is: 1 portuguese, 1 dutch, 1 belgium, 1 finish, 1 hungarian, 2 maltese, 2 british, 2 greek, 2 polish, 3 french, 3 slovenian, 4 cypriot, 5 german, 5 italian, 8 spanish, 37 american. This sums up to 78 SA banks. The selection criteria is the same, asset size.

```{r summary, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
bank.data %>% 
  subset(select = c("savings_log","PD_mean","gini","Tier.1.Ratio","ROE.using.P.L.before.tax....","Z-score","asset_log","tier.1_log")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("log Savings","Portfolio mean PD","Gini coefficient","CT1 ratio","ROE","Z-score", "log Total asset","log CT1 capital"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{11cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

# Empirical strategy
In this section I briefly describe the empirical strategy of the paper.

## EBA capital exercise and the effect on capital savings

To find out if banks adjust their portfolio risk distribution in response to capital requirement shock, I intend to explore the 2011 EBA capital exercise. This was a strong and unexpected test to the solvency of 63 european banks that resulted in a recommendation for these banks to build up their capital buffert to reach a 9.0% core tier 1 capital ratio by June 2012. Identification of the effect of higher capital requirements on banks'portfolio risk distribution comes from comparing the change in the _Capital savings_ measure (or other measure of the distribution such as the Gini coefficient) around the exercise between banks that participated in the exercise and banks that did not. In particular, I (will) estimate the following specification::

\begin{equation}
\Delta\text{Capital savings}_{i}= \beta\text{CEB}_{i} + \varepsilon_{i},
\end{equation}

where $\text{CEB}_{i}$ is an indicator function that equals one if bank $i$ was included in the exercise and $\beta$ is the treatment effect of higher requirements on portfolio risk distribution.

##  Basel II and the effect of capital savings

Once we know that banks adjust their portolio risk distribution to achieve regulatory capital requirements, an natural extension is to quantify the effect of these changes on capital ratio and evaluate if these changes impact other variables of interest, for instance, return on equity and risk-taking. Endogeneity is an important issue to consider. For instance, worse capitalized banks have a stronger incentive to adjust their portfolios compared to their better capitalized peers resulting in downward biased estimates. Conversely, better-capitalized banks may have more leeway to achieve a portfolio distribution that maximizes return by increasing risk-taking. In this case the estimated effects of capital savings on return on equity and risk-taking will have a positive bias.

To solve the endogeneity concerns and estimate the causal effect of portfolio risk distribution, I (will) explore the time differences in introduction of Basel II across countries. Only after Basel II that the IRB models were allowed to be used and, consequently, the distribution of PD matter for the calculation of RWA. In particular, I (will) estimate the following specification:

\begin{equation}
Y_{it}= \beta\widehat{\text{Capital savings}}_{it}+\varepsilon_{i},
\end{equation}

where $Y_{it} \in \{\text{CT1 ratio},\text{ROE},\text{Z-score}\}$ and $\widehat{\text{Capital savings}}_{it}$ is the fitted values of the first stage equation:

\begin{equation}
\text{Capital savings}_{it}= \alpha\text{Basel II}_{it}+\epsilon_{i},
\end{equation}

where $\text{Basel II}_{it}$ is an indicator function that equals one if bank $i$ in year $t$ is hosted in country that has adopted Basel II and has its own IRB model approved.

# (Preliminary) Results

reg.hdfe1 <- felm(Tier.1.Ratio  ~ savings_log  | 0  |0|0,
                  data = bank.data)

reg.hdfe1.1 <- felm(Tier.1.Ratio  ~ savings_log + asset_log + tier.1_log +PD_mean | 0  |0|0,
                  data = bank.data)

reg.hdfe1.2 <- felm(Tier.1.Ratio  ~ savings_log + asset_log + tier.1_log +PD_mean | year +name  |0|0,
                  data = bank.data)

reg.hdfe2 <- felm(ROE.using.P.L.before.tax....~ savings_log  | 0 |0|0,
                  data = bank.data)
reg.hdfe2.1 <- felm(ROE.using.P.L.before.tax....~ savings_log + asset_log + tier.1_log +PD_mean | 0 |0|0,
                  data = bank.data)
reg.hdfe2.2 <- felm(ROE.using.P.L.before.tax....~ savings_log + asset_log + tier.1_log +PD_mean | year +name |0|0,
                  data = bank.data)

>>>>

library(tidyverse)           ## Data manipulation, pipe operator
library(splitstackshape)     ## Command to expand data
library(data.table)          ## Command to bind lists
library(ggridges)            ## Density ridges plot
library(stargazer)           ## Latex tables
library(lfe)                 ## High-dimensional fixed effects
library(gridExtra)           ## Plot side-by-side
library(DescTools)           ## Command to winsorize variables
library(robustHD)

p_w <- c(0.01, 0.99)         ## Set winsorizing limits
```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
load("../../Data/Datasets/DifInDif.Rda")
```

@@

test <- cross.section.decomposition %>% mutate(mean.total.gain = mean(total.gain,na.rm = TRUE)) %>%
  mutate_at(vars(SA.gain:total.gain),  funs((. - mean.total.gain)^2))%>%
summarize_at(vars(SA.gain:total.gain),  sum, na.rm =  TRUE)%>%
  mutate_all( funs((. / total.gain)))
```

```{r}

bank.data <- bank.data %>% ungroup() %>%
  mutate(RW.hat=RW.hat*100,
         RW.s=RW.s*100)
```

# List
* Use the share under IRB to weight capital savings measure (share vs intensity). Include share in regressions plus interaction with capital savings

* Explore alternative capital savings measures, such as, savings in terms of capital (mn),in terms of capital ratio, in terms of capital ratio savings (counterfactual vs actual)

* Plot marginal increase of savings in time.

* Decompose capital savings: share effect, size effect, etc...

* Remove step effect from regressions: average across all IRB banks (add IRB dummy) or IRB specific dummy (average out the step for each bank). Idea is to focus on difference across IRB banks. it might not make sense to include SA banks anymore.

* Plot capital savings measures to other risk variables

* Plot cross section for different capital saving measures (replicate first plot)

* check and better clean data:

# Capital savings measure

@@

\begin{equation}
\text{Capital savings}_{i}= \frac{f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i)}{\sum_{i}EAD_i},
\end{equation}

@@

```{r}
savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
```

```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
 ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
                          y = RW.s,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital savings") +

@@

```{r, fig.height = 10, fig.width = 7, eval=FALSE, include=FALSE, results="hide",fig.cap= fig3_cap,out.width = "1\\textwidth"}
pillar3.expanded %>% 
  filter( !is.na(PD),PD<100, freq_type%in%c("EL") ) %>% 
ggplot(aes(x = PD, y = reorder(name, PD,FUN = median, na.rm=TRUE),color=Country,fill=Country)) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.6,
                       scale = 2 , from = 0, to = 0.05) +
  scale_x_continuous(expand = c(0.01, 0),labels = scales::percent) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(
    x = 'Probability of default') +
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())
```

# Decomposition

\begin{equation}
\begin{aligned}
\Delta RW & = q_{B}^{SA} \Delta RW^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \hat{RW}^{j} \\ 
& + (1-q_{B}^{SA})\sum_{j=1}^{P} \hat{RW}^{j} \Delta q^{j} -(1-q_{B}^{SA})\Delta RW^{S} + (RW^{SA}-RW^{IRB})\Delta q^{SA},
\end{aligned}
\end{equation}

\newpage
## Time decomposition

```{r}
time.dec_cap = "Time decompostion. The figure shows the contribution of each component compared to the time of IRB introduction (t=0). The figure and calculations include only banks with observations for at least 5 consecutive years. The black line is the net change. Year decompostion. The figure shows the contribution of each component compared to 2008. The figure and calculations include all banks in the sample. The black line is the net change." 
```

```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
time.lable <- c("t=0", "t=1", "t=2", "t=3", "t=4")
plot1 <-time.decomposition %>% 
  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain"),
         RW = RW - t0.RW,
         time = as.factor(time)) %>%
  ggplot(mapping = aes(x = time,
                       y    = value, 
                       fill = condition)) +
geom_bar(stat = 'identity')+
geom_line(aes(group=1, y = RW))  +
  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
  scale_x_discrete( name=NULL,breaks = c("0","1","2","3","4"),labels = c("t=0","t=1","t=2","t=3","t=4")) +
 theme(legend.position = c(0.3, 0.2), 
        legend.title = element_blank(), 
        legend.background=element_blank())

year.decomposition <- year.decomposition %>%
  mutate(RW_2 = RW - t0.RW,
         year = as.factor(year)) 
plot2 <- year.decomposition %>% 
  select(year,RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = year)) +
geom_bar(aes(y = value, 
          fill = condition),
          stat = 'identity')+
geom_line(data = year.decomposition, 
          aes(group=1,y = RW_2))  +
  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
theme(legend.position='none')
grid.arrange(plot1, plot2, ncol=2)

\newpage

## Cross-section decomposition

```{r}
cross.dec_cap = "Cross-section decompostion. The figure shows the contribution of each component compared to the average bank. The figure and calculations include all banks in the sample. The black line is the total RW. Relative cross-section decompostion. The figure shows the relative contribution of each component compared to the average bank. The figure and calculations include all banks in the sample. " 
```

```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
plot1 <- cross.section.decomposition %>% 
  mutate(order.s = abs(dispersion.gain))%>%
  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = reorder(name, order.s),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  geom_line(aes(y = total,group=1))  +
  labs(x=NULL, y= NULL,subtitle = "Plot a") +
    coord_flip() +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
 theme(legend.position='none')

plot2 <- cross.section.decomposition %>% 
  mutate(order.s = abs(dispersion.gain))%>%
  select(name, Country, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(value = abs(value),
         condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = reorder(name, order.s),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity',position = "fill")+
  labs(x=NULL, y= NULL,subtitle = "Plot b") +
    coord_flip() +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
   theme(legend.title = element_blank(), 
        legend.background=element_blank(),
       legend.text=element_text(size=7),
       axis.text.y = element_blank())

grid.arrange(plot1, plot2, ncol=2)

# Counterfactual analysis

```{r}
tier1_cap = "The figure shows scatter plots and the linear relationship between the capital savings measure and capital ratios. The y-axis variable on plots a and b is the tier 1 capital ratio, and on plots c and d is the total capital ratio. Plots a and c plots all data points in the sample. Plots b and d plots the average across time for bank using the IRB approach."
```

```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
 plot1 <- bank.data %>%  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>% 
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
theme(legend.position='none')

plot2 <- bank.data %>%  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="",subtitle = "Plot b") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  theme(legend.position = c(0.8, 0.85), 
        legend.title = element_blank(), 
        legend.background=element_blank())

plot3 <- bank.data %>%  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

plot4 <- bank.data %>% filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="",  subtitle = "Plot d") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
```

```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  subset(select = c("RW.s","tier1ratio","tier.hat","totalcapitalratio","CAR.hat","ROE","Zscore")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("Capital savings","Tier 1 ratio","Counterfactual tier 1 ratio","Total ratio","Counterfactual total ratio","ROE","Z-score"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{11cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

```{r}
fig2_cap = "The figure plots the evolution in time of Capital savings, Tier 1 capital ratio and, ROE. The solid line is the average across banks for each year. The first and third quartiles are shown in the gray area."
```

```{r, fig.height = 3, fig.width = 7, results="hide",fig.cap= fig2_cap,out.width = "1\\textwidth", eval=FALSE, include=FALSE}
mean_cl_quantile <- function(x, q = c(0.25, 0.75), na.rm = TRUE){
  dat <- data.frame(y = mean(x, na.rm = TRUE),
                    ymin = quantile(x, probs = q[1], na.rm = na.rm),
                    ymax = quantile(x, probs = q[2], na.rm = na.rm))
  return(dat)
}

plot1 <- bank.data %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu) %>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>% 
ggplot(aes(x=time)) +
 geom_smooth(aes( y=RW.s, color = RW.s),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Capital savings') +
  theme_bw()

plot2 <- bank.data  %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu) %>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=ROE, color = ROE),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'ROE') +
  theme_bw()

plot3 <- bank.data  %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  mutate(time = lubridate::ymd(year, truncated = 2L))%>%
ggplot(aes(x=time)) +
   geom_smooth(aes(y=tier1ratio, color = tier1ratio ),stat = 'summary', fun.data = mean_cl_quantile,color="black")+
labs(x="",y="",
     title = 'Tier 1 ratio') +
  theme_bw()

@@

```{r, include=FALSE}

DiD.data <- DiD.data %>% ungroup() %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)

DiD.intensity <- felm(outcome ~ Shortfall.RW | Country |0|0,
                  data = DiD.data)

DiD.intensity.2 <- felm(outcome ~ Shortfall.RW.2 | Country |0|0,
                  data = DiD.data)

DiD.dummy <- felm(outcome ~ EBAbank | Country |0|0,
                  data = subset(DiD.data, subset = EBAcountry == 1))
